A search for intermediate-mass black holes mergers in the \\second LIGO--Virgo observing run with the Bayes Coherence Ratio

### 2021-06-15 images/bcr_formalism.tex
submission to dcc

\documentclass[%
%reprint,
%superscriptaddress,
%groupedaddress,
%unsortedaddress,
%runinaddress,
%frontmatterverbose, 
% preprint,
%preprintnumbers,
 nofootinbib,
%nobibnotes,
%bibnotes,
 amsmath,amssymb,
 aps,
%pra,
%prb,
%rmp,
%prstab,
%prstper,
%floatfix,
 twocolumn,
 superscriptaddress
]{revtex4-2}

\usepackage{aas_macros}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{array}
\usepackage{color,units}
\usepackage[dvipsnames]{xcolor} % for more colours!
\usepackage{lineno}
\usepackage{dcolumn}
\usepackage{longtable}
\usepackage[normalem]{ulem} %% for striking out text
\usepackage{subfigure}
\usepackage[T1]{fontenc}
\usepackage[breaklinks]{hyperref}
\usepackage{booktabs}
\usepackage{xspace} 
\usepackage{bm}% bold math
\usepackage{graphicx} % Include figure files
% \usepackage{authblk}

\graphicspath{{images/}} %Setting the graphicspath

% symbols
\DeclareSymbolFont{extraup}{U}{zavm}{m}{n}
\DeclareMathSymbol{\varheart}{\mathalpha}{extraup}{86}
\DeclareMathSymbol{\vardiamond}{\mathalpha}{extraup}{87}

% Keywords
\newcommand{\bilby}{{\sc \href{https://lscsoft.docs.ligo.org/bilby/}{\texttt{Bilby}}}\xspace}
\newcommand{\bilbypipe}{{\sc \texttt{bilby\_pipe}}\xspace}
\newcommand{\dynesty}{{\sc \texttt{dynesty}}\xspace}
\newcommand{\cpnest}{{\sc cpnest}\xspace}
\newcommand{\ptemcee}{{\sc ptemcee}\xspace}
\newcommand{\gwpy}{{\sc \href{https://gwpy.github.io/}{\texttt{GWpy}}}\xspace}

\newcommand{\imrphenomp}{{\sc \texttt{IMRPhenomPv2}}\xspace}
\newcommand{\seob}{{\sc \texttt{SEOBNRv4PHM}}\xspace}
\newcommand{\nrsur}{{\sc \texttt{NRSur7dq4}}\xspace}
\newcommand{\imrxhm}{{\sc \texttt{IMRPhenomXPHM}}\xspace}

\newcommand{\gstlal}{{\sc GstLAL}\xspace}
\newcommand{\cwb}{{\sc cWB}\xspace}
\newcommand{\spiir}{{\sc SPIIR}\xspace}
\newcommand{\mbta}{{\sc MBTAOnline}\xspace}
\newcommand{\pycbc}{{\sc \href{https://pycbc.org/}{{PyCBC}}}\xspace}

\newcommand{\GWTC}{{\sc \href{https://ui.adsabs.harvard.edu/abs/2019PhRvX...9c1040A/abstract}{{GWTC-1}}}\xspace}
\newcommand{\OGC}{{\sc \href{https://ui.adsabs.harvard.edu/abs/2020ApJ...891..123N/abstract}{{2-OGC}}}\xspace}
\newcommand{\IAS}{{\sc \href{https://ui.adsabs.harvard.edu/abs/2020PhRvD.101h3030V/abstract}{{IAS}}}\xspace}

% math keywords 
\newcommand{\fancytext}[1]{{\relax\ifmmode#1\else $#1$\fi}\xspace}
\newcommand{\mathcmd}[1]{{\sc \relax\ifmmode#1\else $#1$\fi}\xspace}
\newcommand{\bcr}{\mathcmd{\rho_\text{BCR}}}
\newcommand{\bcrodds}{\mathcmd{\mathcal{O}_{\mathrm{BCR}}}}
\newcommand{\pycbcstat}{\mathcmd{\rho_\text{PC}}}
\newcommand{\snr}{\mathcmd{\rho}}
\newcommand{\psd}{\mathcmd{P(f)}}
\newcommand{\msun}{\mathcmd{\text{M}_\odot}}
\newcommand{\mtot}{\mathcmd{M}}
\newcommand{\parameters}{\mathcmd{\vec{\theta}}}
\newcommand{\prior}{\mathcmd{\pi(\parameters)}}
\newcommand{\template}{\mathcmd{\mu(\parameters)}}
\newcommand{\fap}{\mathcmd{\text{FAP}}}

\newcommand{\pastro}{\fancytext{p_\text{astro}}}
\newcommand{\pastrobcr}{\fancytext{p_\text{astro}^{\text{BCR}}}}
\newcommand{\untunedpastrobcr}{\fancytext{p_\text{astro}^{\text{BCR}\prime}}}
\newcommand{\pastroext}{\fancytext{p_\text{astro}^{\text{ext}}}}
\newcommand{\pval}{\fancytext{\text{p-value}}}
% Table settings
\renewcommand{\aboverulesep}{0pt}
\renewcommand{\belowrulesep}{0pt}
% \setlength\cellspacetoplimit{2pt}
% \setlength\cellspacebottomlimit{2pt}

\newcolumntype{C}{>{\centering\arraybackslash}X}
\newcolumntype{L}{>{\arraybackslash}X}

% Aesthetic colours
\definecolor{dodgerblue}{HTML}{1E90FF}
\definecolor{viennared}{HTML}{DA0A14}
\definecolor{ctorange}{HTML}{FF6C0C}
\definecolor{wales}{HTML}{ff0038}
\definecolor{benettongreen}{HTML}{009421}
\definecolor{valenciacfred}{HTML}{ee3524}
\definecolor{barcelonafcgold}{HTML}{edbb00}
\definecolor{jam}{HTML}{A50B5E}
\definecolor{austriawien}{HTML}{441678}
\definecolor{italia90green}{HTML}{009966}
\definecolor{ferrarired}{HTML}{ff2800}
\definecolor{gray}{HTML}{F0F0F0}
\definecolor{LightCyan}{rgb}{0.88,1,1}
\newcolumntype{a}{>{\columncolor{gray}}c}
\newcolumntype{b}{>{\columncolor{white}}c}

% author comments
\newcommand{\avi}[1]{\textcolor{orange}{[AV: #1]}}
\newcommand{\greg}[1]{\textcolor{purple}{[Greg: #1]}}
\newcommand{\rs}[1]{\textcolor{red}{[RS: #1]}}
\newcommand{\et}[1]{\textcolor{blue}{[ET: #1]}}

\begin{document}

\title{A search for intermediate mass black holes mergers in the \\second LIGO--Virgo observing run with the Bayes Coherence Ratio}

\input{authors}

\date{\today}

\begin{abstract}
The detection of an intermediate-mass black hole population ($10^2-10^6$~\msun) will provide clues to their formation environments (e.g., disks of active galactic nuclei, globular clusters) and illuminate a potential pathway to produce supermassive black holes. Ground-based gravitational-wave detectors are in principle sensitive to such mergers and have been used to detect one $142^{+28}_{-16}\ \msun$ intermediate-mass black hole formation event. Ground-based detector data contain numerous short-duration noise transients that can mimic the gravitational-wave signals from merging intermediate-mass black holes, limiting the sensitivity of searches. Here we demonstrate a Bayesian-inspired ranking statistic to detect binary black hole mergers with a total mass $\gtrsim55$~\msun. We use this statistic to identify candidate events with total masses $>55$~\msun using data from LIGO's second observing run. Our analysis does not yield evidence for new intermediate-mass black holes. However, we find support for some stellar-mass binary black holes not reported in the first LIGO--Virgo gravitational-wave transient catalog, GWTC-1.
\end{abstract}

% \avi{Say why it improves on prev techniques}

\maketitle

%%%%%%---SECTIONS---%%%%%%%%%%%%
\section{Introduction}
Since the 1970s, there has been a steady accumulation of evidence for stellar mass ($M_\text{BH} < 10^{2} \ \msun$) and supermassive black holes ($M_\text{BH} > 10^{6} \ \msun$)~\cite{Webster:1972:Natur, Balick:1974:ApJ, Ghez:1998:ApJ, Genzel:2010:RvMP, Abbott:2019:PhRvX, EventHorizonTelescopeCollaboration:2019:ApJL, Abbott:2020:arXiv}.  However, there is a deficiency of observational evidence for black holes in the intermediate-mass range $10^{2} - 10^{6}\ \msun$. The discovery of an intermediate-mass black hole population will bridge this observational gap, probe intermediate-mass formation environments (e.g. accretion disks of active galactic nuclei~\cite{Tagawa:2021:ApJ, Li:2021:arXiv, Samsing:2020:arXiv, Tagawa:2020:ApJ, Ishibashi:2020:A&A, Grobner:2020:A&A, Yang:2019:PhRvL, McKernan:2019:ApJL, Yang:2019:ApJ, McKernan:2018:ApJ, Bellovary:2016:ApJL, McKernan:2014:MNRAS, McKernan:2012:MNRAS}, the centers of dense stellar clusters~\cite{Banerjee:2021:MNRASa, Zevin:2021:ApJ,Mapelli:2021:arXiv,Weatherford:2021:ApJL, Bouffanais:2021:arXiv, Ballone:2021:MNRAS, Kumamoto:2021:arXiv, Banerjee:2021:MNRASb, Martinez:2020:ApJ, Romero-Shaw:2020:ApJL, Anagnostou:2020:PASA}, Population-III stars~\cite{Toubiana:2021:PhRvL, Farrell:2021:MNRAS, Safarzadeh:2020:ApJL, Liu:2020:MNRAS, Inayoshi:2017:MNRAS}), and illuminate our understanding of supermassive black hole formation~\cite{Askar:2021:MNRAS, ArcaSedda:2019:arXiv, Amaro-Seoane:2007:CQGra, Gurkan:2006:ApJL}.

A variety of techniques have been employed to search for $10^{4} - 10^{6}\ \msun$ intermediate mass candidates including reverberation mapping~\cite{Peterson:2014:SSRv}, direct kinematic measurements~\cite{Schodel:2002:Natur, Kiziltan:2017:Natur}, applying macroscopic galaxy to black hole mass scaling relations ($M_{BH}$-$\sigma$ and $M_{BH}$-L relations)~\cite{Graham:2013:ApJ, Wevers:2017:MNRAS}, studying  X-ray luminosity and spectra~\cite{Greene:2004:ApJ, Lin:2020:ApJL}, gravitational lensing of gamma-ray burst light curves~\cite{paynter_evidence_2021}, and others~\cite{Greene:2020:ARA&A, Koliopanos:2017:mbhe, Mezcua:2017:IJMPD}). However, these observational techniques are challenging to use for intermediate mass black holes due to their small sphere of influence compared to super-massive black holes~\cite{Mezcua:2017:IJMPD}. Additionally, some IMBH candidates can be attributed to sources other than intermediate-mass black holes (e.g., clusters of stellar-mass black holes~\cite{Ridolfi:2016:MNRAS, Freire:2017:MNRAS}, anisotropic emission from neutron stars~\cite{Israel:2017:MNRAS, RodriguezCastillo:2020:ApJ}), while others have high uncertainties intrinsic to the observational techniques~\cite{Greene:2020:ARA&A}. 
\et{Consider replacing previous sentence that says none of the current IMBH candidates are widely accepted within the astronomical community as unambiguous.}

Compact binary coalescences (CBCs) can provide unambiguous gravitational-wave signals for intermediate-mass candidates (e.g., the $142^{+28}_{-16}\ \msun$ remnant observed from the gravitational wave event GW190521~\cite{Abbott:2020:PhRvL}). As a binary's total mass $\text{M}_{T}$ is associated with its gravitational-wave merger frequency, $f\sim \text{M}_{T}^{-1}$,  ground-based gravitational-wave detectors ($f\sim 10^1 - 10^3\ \text{Hz}$) are sensitive to the last milliseconds of merging systems with $100 \msun < \text{M}_{T} < 400 \msun$~\cite{LIGOScientificCollaboration:2015:CQGra, Martynov:2016:PhRvD, Moore_2014}, while space-based detectors ($f \sim 10^{-2} - 10^1\ \text{Hz}$) can study the full signals of merging systems with $10^4 \msun < \text{M}_{T} < 10^7 \msun$~\cite{ Moore_2014, Lu:2019:PhRvD}. Because of the short-duration of intermediate-mass gravitational-wave signals in ground-based detectors, data quality is critical for their detection. Gravitational-wave data is characterized by numerous non-stationary terrestrial artifacts called \textit{glitches}~\cite{ pycbc_short_duration_transients, pe_with_glitch, blip_glitches}. Like signals from intermediate mass mergers, most glitches last for a fraction of a second, making them difficult to distinguish from astrophysical signals. These glitches can decrease the sensitivity of searches for binary black hole mergers with total masses $>55$~\msun ~\cite{pycbc_short_duration_transients}.

Although a significant fraction of the glitches can be identified by testing them for coherence amongst two or more detectors and performing matched-filtering, these methods are insufficient to identify all glitches~\cite{ pycbc_short_duration_transients, pe_with_glitch, blip_glitches}. One method to discriminate more glitches while searching for CBC is the Bayesian odds~\cite{bci, kanner2016leveraging, BCR1, BCR2, bcr_gw151216, bayesian_odds}. \et{Need sentence here saying we are inspired by the Bayesian odds to develop a ranking statistic that incorporates some of the ideas used in the odds.} In this paper, we rank O2's candidate gravitational-wave signals from $55-500\ \msun$ (\et{lab frame}) total mass systems using a ranking statistic called the Bayesian Coherence Ratio \bcr~\cite{BCR1}. \et{I deleted some text here going into detail about the BCR; save for method section.} For each candidate, we calculate the probability of astrophysical origin, \pastrobcr and compare to candidate events reported by other authors including the LIGO-Virgo-KAGRA (LVK) collaboration in \GWTC~\cite{GWTC1}, the \pycbc-team~\cite{pycbc_code, pycbc_og0, pycbc_og1, pycbc_og2, pycbc_og3, pycbc_og4, pycbc_og5, pycbc_og6, pycbc_single_det, pycbc_ogc_2}, by the Institute of Advanced study's team (\IAS)~\cite{IAS0, IAS1, IAS2}, or by \citet{bayesian_odds}. \et{Deleted several sentences here about innovations that I think are in the weeds.}

We find that (a) high-mass events reported in the \GWTC, including GW170729 (the heaviest event in \GWTC) are very statistically significant $\pastrobcr>0.9$; (b) three out of the eight \IAS events and candidates have $\pastrobcr>0.5$, corroborating \IAS's detection claims for GW170304, GW170727, and GW170817A; and that (c) our ranking statistic does not identify any new intermediate-mass black holes, but does identify an unreported marginal stellar-mass binary black hole candidate, 170222 with $\pastrobcr\sim0.5$. %\avi{Need to doublecheck that this wasnt in the PyCBC sub-threshold paper}.

The remainder of this paper is structured as follows. We outline our methods, including details of our ranking statistic and the retrieval of our candidates in Section~\ref{sec:method}. We present details on the implementation of our analysis in Section~\ref{sec:Analysis}. Finally, we present our results in Section~\ref{sec:Results}, and discuss these results in the context of the significance of gravitational-wave candidates in Section~\ref{sec:Conclusion}.

\section{Method\label{sec:method}}

The standard framework to identify CBC gravitational-wave signals in data is to quantify the significance of candidates with null-hypothesis significance testing~\cite{GWTC1, GWTC2}. In this framework, the candidates' ranking statistic is compared against a background distribution. The independent matched-filter searches, e.g., \pycbc~\cite{pycbc_og4}, \spiir~\cite{spiir} and \gstlal~\cite{sachdev2019gstlal}, and Coherent WaveBurst~\cite{cwb} used by LVK to search for signals in gravitational-wave data all use ranking statistics in such a manner~\cite{GWTC1}. Both \pycbc and \gstlal's ranking statistic incorporate information about the relative likelihood that the data contains a coherent signal versus noise. In contrast, \cwb's ranking statistic uses the information of coherent energy present in the network of detectors~\cite{GWTC1}.

Bayesian inference offers an alternative means to rank the significance of candidate events by computing the odds that the data contain a transient gravitational-wave signal versus instrumental glitches~\cite{BCR1}. This method relies on accurate models for the signal and glitch morphologies~\cite{BCR1}. In principle, Bayesian odds is the optimal method for hypothesis testing~\cite{BCR2}. Much of its power comes from the Bayesian evidence, the likelihood of the data given a hypothesis. However, the evidence is not used in current matched filter searches. Here, we explore a hybrid frequentist/Bayesian ranking statistic that makes use of the Bayesian evidence. We compute the Bayesian evidence under the assumption that they either contain a coherent gravitational-wave signal, noise, or a glitch. However, instead of computing true Bayesian odds, we use the evidences as a ranking statistic.
We form a bootstrapped distribution of the evidence for simulated foreground and background events in order to form a frequentist ranking statistic.

@@

The \bcrodds quantity is a Bayesian odds-ratio like the above, of a coherent signal hypotheses $\mathcal{H}_S$ and an incoherent instrumental feature hypothesis $\mathcal{H}_I$ (the null-hypotheses) for a network of $D$ detectors. $\mathcal{H_I}$ states that each detector $i$ has either pure stationary Gaussian noise $\mathcal{H}_N$ or Gaussian noise and an incoherent noise transient (glitch) $\mathcal{H}_G$. Taking $Z^S$, $Z^G_i$ and $Z^N_i$ as the Bayesian evidences (defined in Appendix~\ref{sec:bayesianEvidEval}) for $\mathcal{H}_S$, $\mathcal{H}_N$, and $\mathcal{H}_G$, \bcrodds is given by
\begin{equation}
\label{eq:bcr}
\bcrodds = \frac{\pi^S Z^S}{\prod\limits^D_{i=1} \ [\pi^G Z^G_i + (1-\pi^G)Z^N_i]}\ ,
\end{equation}

@@

When $\mathcal{H}_S$ and $\mathcal{H}_I$ are precisely described and the prior-odds represent our true beliefs, the \bcrodds is a Bayesian odds-ratio. As an odds-ratio, the \bcrodds is the optimal discriminator between coherent signals and incoherent instrumental features. However, as the priors-odds are unknown, we \textit{tune} values for prior-odds, $\hat{\pi}^S$ and $\hat{\pi}^G$, to estimate \bcrodds with a ranking statistic, \bcr, given by
\et{I would remove everything in this subsection before this point. Save it for your dissertation if you want, but I don't think we need to review Bayesian basics here. Instead, just begin by saying that we implement the BCR, which is given by Eq.~\ref{eq:rho_bcr} with suitable references to the previous BCR papers. The motivation for the BCR can be summarised in a single sentence.}
\begin{equation}
\label{eq:rho_bcr}
\bcr = \frac{\hat{\pi}^S Z^S}{\prod\limits^D_{i=1} \ [\hat{\pi}^G Z^G_i + (1-\hat{\pi}^G)Z^N_i]}\ .
\end{equation}
\et{Rewrite the remaining paragraphs to explain that how the BCR relates to the odds with just a few sentences, avoiding equations.}
In the limit where the estimated prior-odds equal the true prior-odds, $\bcr\to\bcrodds$. However, as we are uncertain what the true prior-odds are, it is invalid to use the \bcr as an odds-ratio to make an informed decision about whether a candidate is from an astrophysical or terrestrial source. Instead of interpreting the \bcr as a Bayesian odds-ratio, it can be used as a ranking statistic. Using the \bcr as a ranking statistic we can obtain a frequentist significance of a candidate \bcr-value measured against a background \bcr distribution.

@@

\subsection{Estimation of astrophysical signal probability}
\et{Above, I propose to cut the previous subsection way back. I think this subsection can be cut way back too and combined with the previous subsection. Once we have finished talking about $\rho_\text{BCR}$, we can just state that we use time-shifted data to generate distributions of $\rho_\text{BCR}$ for noise, which we use to calculate a single-event false alarm probability = the fraction of triggers with a $\rho_\text{BCR}$ at least that big. (The thing we are currently calling $f$ is a single-event false alarm probability, so we should call it that and give it a better variable name like $p_1$.) Then comment that we consider $N$ triggers so we also calculate a false-alarm probability with trial factors, which is given by the current Eq.~6. (I suggest you call the false alarm probability with trial factors $p_N$ to avoid confusion.) The text should say how $N$ is determined. I agree that $p_\text{astro} = 1-p_N$ is too sloppy to do in a paper. (The reason we are using FAP is because there was not appetite to calculate the distribution of $\rho_\text{BCR}$ for a plausible signal distribution, which is necessary to get an actual $p_\text{astro}$.) At this late stage, I think the best thing to do is just say that we are going to report $1-p_N$ alongside $p_\text{astro}$ values even though they are apples and oranges. While it's sloppy to compare them directly, they \textit{are} both indicators of significance. Moreover, different groups make different assumptions about the signal distribution, and so the different $p_\text{astro}$ values are sort of already different fruit; see Shanika's paper. We can explain all this and say we're going to put everything into one table anyway. However, the least bad thing to do is to report $p_N$ without calling it $p_\text{astro}$. I think most referees can live with that.}
Candidate \bcr-values are either statistically insignificant compared to the background \bcr distribution, implying the candidate is more probable to be an incoherent instrumental feature (the $\mathcal{H}_I$ null-hypothesis), or statistically significant relative to the background distribution, indicating the possible presence of an astrophysical signal (the $\mathcal{H}_S$ hypothesis). A false alarm probability with trial factors, \fap, for the candidate \bcr-value can quantify the significance. The \fap for a \bcr-value is the probability that a candidate originating from a non-astrophysical source can be falsely identified as a signal once in $N$ trials, and is given by 
\begin{equation}
    \fap = 1 - (1-f)^N \ ,
\end{equation}
where $f$ is the probability of observing a background $\bcr'$ greater than or equal to the candidate \bcr,
\begin{equation}
    f = \frac{\text{Count of } \bcr' \leq \bcr}{\text{Count of } \bcr'} \ .
\end{equation}

Finally, the \fap can be used to construct a \pastro, the probability that a signal is of astrophysical origin~\cite{pastro_1,pastro_2,pastro_3}
\begin{equation}
    \pastro = 1 -  \fap \ .
\end{equation}

\avi{Eric, can you take a look at this section about the FAP p-astro?}
\avi{
The pastro calculation may need some more discussion. Taking pastro = 1-FAP as identifying a real signal can be hugely problematic: \url{https://en.wikipedia.org/wiki/Misuse_of_p-values}, even though its ok in this case. Maybe the following papers have something that can help motivate this ~\cite{Farr:2015:PhRvD, Gaebel:2019:MNRAS,Galaudage:2020:PhRvD}
}

\subsection{Background estimation}

To rank candidate event \bcr values we compute a background \bcr distribution. We use time-shifted background data from \pycbc~\cite{pycbc_og4}. \et{Add sentence here explaining what time-shifting is and why it is used.} The output of \pycbc's search is a list of times and their corresponding \pycbc ranking statistic \pycbcstat values. Whenever a local maximum of $\pycbcstat > \snr_\text{T}$, where $\snr_\text{T}$ is some predetermined threshold value, the \pycbc search pipeline produces a single-detector \textit{trigger} associated with the detector and time $t_c$ where the apparent signal in the data has its merger~\cite{pycbc_og6}.
For each \pycbc background trigger, we calculate \bcr.
% \avi{Should I mention that we need \pycbc  as a pre-selection step to computing? }
\et{I think this section can be cut way back too. (I've deleted a bunch of stuff w track changes.) I don't think we need to get into the three types of triggers; just focus on background triggers. I would move this discussion into a single subsection along with IIA and IIB. I would cover this stuff \textit{before}getting into FAP calculation. That way, we already have introduced the distribution of $\rho_\text{BCR}$; it will be more clear what we are talking about.}

\section{Analysis}\label{sec:Analysis}

\subsection{Acquisition of triggers}
Advanced LIGO's second observing run O2 lasted $38$ weeks~\cite{GWOSC}. The software package, \pycbc~\cite{pycbc_code}, was used by LVK to process the O2 data in 22 time-frames (approximately 2 weeks per frame) and found several gravitational-wave events and numerous gravitational-wave candidates~\cite{pycbc_og0, pycbc_og1, pycbc_og2, pycbc_og3, pycbc_og4, pycbc_og5, pycbc_og6}. Some candidate events were vetoed to be glitches, while others were rejected due to their low significance. The data are divided into these time-frames because the detector's sensitivity does not stay constant throughout the eight-month-long observing period.

In addition to finding candidate events, \pycbc also identified several million background triggers for each time-frame, by searching background data manufactured by time-sliding data within that time-frame. The background triggers help quantify the candidate events' significance for the respective time-frames. Finally, to test the search's sensitivity, \pycbc produced and searched for thousands of simulated signals. 
% \avi{Do i need to contextualize this choice more, ie talk about what would happen if I chose a narrower range?}

For our study, we filter the \pycbc background, simulated and candidate triggers to include only triggers in the ranges of the parameters presented in Table~\ref{tab:parameters}. Filtering triggers to this region focuses our analysis to binary black hole mergers with total masses above $> 55 \msun$. This corresponds to binary systems with signal durations $<454 \ \text{ms}$, signals which may be mistaken for short-duration glitches. A plot of the \pycbc triggers from one time-frame, during April 23 - May 8, 2017, is presented in Fig.~\ref{fig:templateBank}. This figure also depicts the gravitational-wave templates used by \pycbc's search through this time-frame of data.

\begin{table}[t]

\caption[BBH parameters correspond to duration $<454\ \text{ms}$]{\label{tab:parameters}Trigger-selection parameter space (parameters correspond to signals with durations $<454 \ \text{ms}$). }
\centering
\begin{tabular}{lrr}
\toprule
  & Minimum & Maximum\\
\midrule
Component Mass 1 [\msun] & 31.54 & 491.68\\
Component Mass 2 [\msun] & 1.32 & 121.01\\
Total Mass [\msun] & 56.93 & 496.72\\
Chirp Mass [\msun] & 8.00 & 174.56\\
Mass Ratio & 0.01 & 0.98\\
\end{tabular}
\end{table}

\begin{figure*}[!ht]

{\centering \includegraphics[width=0.75\linewidth]{images/template_bank.png}

}
% time for chunk 14
% 1176955218, Apr 23, 2017, 4:00 UTC
% 1178294418, May 08, 2017, 16:00 UTC
\caption[ BCR search space.]{The templates (pink) used by \pycbc to search a section of O2 data from $\text{April 23 - May 8, 2017}$. Our search is constrained to the parameter space enclosed by the dashed line. The candidate, background and simulated triggers detected in this region of the parameter space during this period are plotted in orange, gray and blue respectively.}\label{fig:templateBank}
\end{figure*}

\subsection{Calculating the BCR for triggers}
To evaluate $Z^S$, $Z^G_i$ and $Z^N_i$ and calculate the \bcr Eq.~\ref{eq:bcr} for triggers, we carry out Bayesian inference with \bilby~\cite{bilby, bilby_pipe}, employing \dynesty~\cite{dynesty} as our nested sampler. Nested sampling, an algorithm introduced by~\citet{skilling2004, skilling2006}, provides an estimate of the Bayesian evidence and is often utilized for parameter estimation within the LIGO collaboration~\cite{bilby, bilby_paper, pbilby_paper}.

The most computationally intensive step during Bayesian inference is evaluating the likelihood $\mathcal{L}(d_i|\template)$. To accelerate our analysis, we use a likelihood that explicitly marginalizes over coalescence time, phase at coalescence, and luminosity distance (Eq.~80 from~\citet{intro_to_gw_bayes}).
\et{I think that IIIa and IIb should probably be shrunken and merged so that Section III does not have any subsections (IIC can be deleted; see below). The way to cut these sections down is to stick to just \textit{describing what was done} rather than trying to add a lot of explanation. (Explaining what PyCBC does and why is particularly excessive, IMO.) Just cut out anything that is not part of a recipe to reproduce the analysis.}

We set the priors $\pi(\parameters|\mathcal{H}_S)$ and $\pi(\parameters|\mathcal{H}_G)$ to be identical which reflects our ignorance of the distribution of the population properties of signals and signal-like glitches. These priors restrict signals with mass parameters in the ranges presented in Table~\ref{tab:parameters}. The spins are aligned over a uniform range for the dimensionless spin magnitude from $\left[0,1\right]$. The luminosity distance prior assigns probability uniformly in comoving volume, with an upper cutoff of $5\ \text{Gpc}$. The full list of the priors, along with their shapes, limits and boundary conditions are documented in Table~\ref{tab:priors}.

\begin{table}
    \centering
    \caption{
    Prior settings for the parameters used during our parameter estimation. The definitions of the parameters are documented in \citet{bilby_gwtc}~Table~E1. The trigger time $t_c$ is obtained from the data products of \pycbc's O2 search. \label{tab:priors}} 
    \begin{tabular}{c c c c}
    \hline
    Parameter & Shape & Limits \\
    \hline
          $\mathcal{M}/\msun$           & Uniform & 7--180  \\
          $q$                           & Uniform & 0.1--1  \\
          $M/\msun$                     & Constraint & 50--500  \\
          $d_\mathrm{L}/\mathrm{Mpc}$   & Comoving & 100--5000  \\
          $a_1$, $a_2$                  & Uniform & 0--1  \\
          $\theta_{JN}$                 & Sinusoidal & 0--$\pi$  \\
          $\psi$                        & Uniform & 0--$\pi$  \\
          $\phi$                        & Uniform & 0--$2\pi$  \\
          ra                            & Uniform & 0--$2\pi$  \\
          dec                           & Cosine & 0--$2\pi$  \\
          $t_c/\mathrm{s}$              & Uniform & $t_c\pm0.1$  \\
    \hline
    \end{tabular}
\end{table}

The waveform template we utilize is \imrphenomp, a phenomenological waveform template constructed in the frequency domain that models the in-spiral, merger, and ring-down (IMR) of a compact binary coalescence~\citep{khan2016frequency}. Although there exist gravitational-wave templates such as \imrxhm~\cite{imrphenompxhm}, \nrsur~\cite{nrsur7dq4} and \seob~\cite{seobnrv4phm} which incorporate more physics, such as information on higher-order modes, we use \imrphenomp as it is computationally inexpensive compared to others.

To generate the PSD, we take 31 neighboring off-source non-overlapping  4-second  segments of time-series data before the analysis data segment $d_i$. To suppress spectral leakage, a Tukey window with a 0.2-second roll-off is applied to each data segment . After this the segments are fast-Fourier transformed and median-averaged to create a PSD~\cite{ligo_psd}. Like other PSD estimation methods, this method adds statistical uncertainties to the PSD~\cite{psd_student_t, chatziioannou2019noise, Biscoveanu:2020:PhRvD}. To marginalize over the statistical uncertainty, we use the median-likelihood presented by~\citet{psd_student_t} as a post-processing step. We find that this post-processing step improves the search efficiency by $49.26\%$ the details of this calculation are presented in the Appendix~\ref{sec:psd-marginalization}.

Finally, we acquire the foreground, background and data in which we inject simulated signals into, from from the gravitational-wave Open Science Center~\cite{GWOSC}. The data we use are the publicly accessible O2 strain data from the Hanford and Livingston detectors, recorded while the detectors are in ``Science Mode''. We obtain the data using \gwpy~\cite{gwpy}.

\subsection{Assigning \pastro to candidate events}
After calculating the \bcr for the entire set of background and simulated triggers, we calculate the background and simulated \bcr probability distributions for each 2-week time-frame of O2 data. These distributions are used to `tune' prior-odd $\hat{\pi}^S$ and $\hat{\pi}^G$ values as described in Appendix~\ref{apdx:tuning-prior-odds}.
\et{This subsection discusses formalism (how the analysis works) and so it feels out of place in Section III, which is about the details of the implementation. I would delete this subsection entirely and just add a sentence or two to the formalism section explaining that the $\pi$'s are tuned emprically; see Appendix~\ref{apdx:tuning-prior-odds}.}

Using the tuned prior-odds the \bcr for the candidate events can be calculated. Fig.~\ref{fig:bcrCdf} shows the \bcr distributions for the background triggers, simulated triggers and candidate events. The bulk of the background and simulated trigger distributions are separate but slightly overlap due to some of the simulated signal's being very faint. The separation suggests that the \bcr can successfully distinguish signals from noise or glitches. The vertical lines in Fig.~\ref{fig:bcrCdf} displays the \bcr for gravitational-wave candidate events. On comparing the candidate event \bcr values with the background distribution, we can estimate \pastro values for the candidate events.

\begin{figure}[!ht]
{\centering \includegraphics[width=0.85\linewidth]{images/reweighted_bcr_cdf_smaller_legend.png} }
\caption[BCR distribution example]{Histograms represent the survival function (1-CDF) from our selection of background triggers (gray) and simulated signals (blue) triggers obtained from \pycbc's search of data from $\text{August 13 - 21, 2017}$. Vertical lines mark the $\text{ln}\ \bcr$ of \IAS's GW170817A and \GWTC's GW170814.}\label{fig:bcrCdf}
\end{figure}

\section{\label{sec:Results}Results}

\input{Data/results_table}
We analyze the $60,996$ background, $5,146$ simulated, and $25$ candidate triggers reported by \pycbc's search on the data from LIGO's second observing run, restricting our analysis to the triggers that fall within our mass-space as described in Section~\ref{sec:method}. We also analyze events and candidate events reported by \GWTC and the \IAS group (note: some of these were identified as candidates by the \pycbc search). Table~\ref{tab:results} summaries the \pastrobcr, along with the \pastro of other pipelines for comparison. 
\et{The previous text about what we analysed should go in Section III. This section is just for reporting the results.}

>>>>

@@

The \bcrodds quantity is a Bayesian odds-ratio like the above, of a coherent signal hypotheses $\mathcal{H}_S$ and an incoherent instrumental feature hypothesis $\mathcal{H}_I$ (the null-hypotheses) for a network of $D$ detectors. $\mathcal{H_I}$ states that each detector $i$ has either pure stationary Gaussian noise $\mathcal{H}_N$ or Gaussian noise and an incoherent noise transient (glitch) $\mathcal{H}_G$. Taking $Z^S$, $Z^G_i$ and $Z^N_i$ as the Bayesian evidences (defined in Appendix~\ref{sec:bayesianEvidEval}) for $\mathcal{H}_S$, $\mathcal{H}_N$, and $\mathcal{H}_G$, \bcrodds is given by
\begin{equation}
\label{eq:origbcr}
\bcrodds = \frac{\pi^S Z^S}{\prod\limits^D_{i=1} \ [\pi^G Z^G_i + (1-\pi^G)Z^N_i]}\ ,
\end{equation}

@@

When $\mathcal{H}_S$ and $\mathcal{H}_I$ are precisely described and the prior-odds represent our true beliefs, the \bcrodds is a Bayesian odds-ratio. As an odds-ratio, the \bcrodds is the optimal discriminator between coherent signals and incoherent instrumental features. However, as the priors-odds are unknown, we \textit{tune} values for prior-odds, $\hat{\pi}^S$ and $\hat{\pi}^G$, to estimate \bcrodds with a ranking statistic, \bcr, given by
\et{I would remove everything in this subsection before this point. Save it for your dissertation if you want, but I don't think we need to review Bayesian basics here. Instead, just begin by saying that we implement the BCR, which is given by Eq.~\ref{eq:bcr} with suitable references to the previous BCR papers. The motivation for the BCR can be summarised in a single sentence.}
\begin{equation}
\label{eq:bcr}
\bcr = \frac{\hat{\pi}^S Z^S}{\prod\limits^D_{i=1} \ [\hat{\pi}^G Z^G_i + (1-\hat{\pi}^G)Z^N_i]}\ .
\end{equation}
\et{Rewrite the remaining paragraphs to explain that how the BCR relates to the odds with just a few sentences, avoiding equations.}
In the limit where the estimated prior-odds equal the true prior-odds, $\bcr\to\bcrodds$. However, as we are uncertain what the true prior-odds are, it is invalid to use the \bcr as an odds-ratio to make an informed decision about whether a candidate is from an astrophysical or terrestrial source. Instead of interpreting the \bcr as a Bayesian odds-ratio, it can be used as a ranking statistic. Using the \bcr as a ranking statistic we can obtain a frequentist significance of a candidate \bcr-value measured against a background \bcr distribution.

@@
